A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-10-04 paper/paper.tex
more

\section{Reinforcement Learning as a computational model for DMN function}
The main idea of this section is to argue that the DMN has biological substrates (gradient
computation, backprop, experience replay, etc.) for all the
major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
components integrate in a plausible way.
Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
environment. In a general RL framework, an \textit{agent}
interacts with an  environment in a trial-by-error fashion by taking
actions / decisions $a$, which trigger changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent. In this view, the environmental is partially controlled by
the action of the agent.
This reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action.
%% Here the subscript $t$ denotes time, and can be discrete or
%% continuous.
The environment $\mathcal E$ is generally taken as stochastic,
is partially observed, in the sense that only part of the current state is observed by
the agent.
\subsection{Markov Decision Processes}
A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
%% In RL, MDPs are the mathematical instantiation of the environment model,
%% which generally can incorporate action probabilities (policy matrix)
%% and/or value function given a state.
%% (Silver: The environment can be fully defined by state space and
%% transition probability matrix.)
is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
\item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$,
  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.  
\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
  $\mu(s)$ is the probability that the environment starts-off in state $s$.
\end{itemize}

>>>>

\section{Reinforcement Learning as a computational model for DMN function}
The main idea of this section is to argue that the DMN has biological substrates (gradient
computation, backprop, experience replay, etc.) for all the
major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
components integrate in a plausible way.
Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
environment. In a general RL framework, an \textit{agent}
interacts with an  environment in a trial-by-error fashion by taking
actions / decisions $a$, which trigger changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent. In this view, the environmental is partially controlled by
the action of the agent.
This reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action.
%% Here the subscript $t$ denotes time, and can be discrete or
%% continuous.
The environment $\mathcal E$ is generally taken as stochastic,
is partially observed, in the sense that only part of the current state is observed by
the agent.
\subsection{Markov Decision Processes}
A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
%% In RL, MDPs are the mathematical instantiation of the environment model,
%% which generally can incorporate action probabilities (policy matrix)
%% and/or value function given a state.
%% (Silver: The environment can be fully defined by state space and
%% transition probability matrix.)
is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
\begin{itemize}
\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
\item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$,
  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.  
\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
  so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
  $\mu(s)$ is the probability that the environment starts-off in state $s$.
\end{itemize}
